---
title: "Incarceration"
author: "Adnan"
date: "11/20/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(maps)
library(mapproj)
library(reshape2)
library(lintr)
source("analysis.R")
read.csv("incarceration_trends.csv")
```


## Introduction
This report was created using the incarceration data set found online. This 
data set shows the incarceration rate over time between 1968 and 2018. It also
shows us the population of each race as well as the jail population in each 
state as well as county. Down below I have created three different charts to 
show visualizations of black jail population, black vs white population in 
each county of Texas, and lastly a map that shows the percentage of latin
people in each county of Texas.


## Summary Information
In my data I tried to focus on race and population of different counties in 
Texas. Therefore I first looked for the state with the highest black jail
population and after calculating it the state with the highest black jail
population was California (```summary_info$state_highest_black_jail_pop```). 
Then I went on to see the county with the highest jail population in the state
of Texas which was Harris County (```summary_info$county_highest_jail_pop_in_TX```).
You can also see in my variable comparison chart that Harris County had the
highest black population in Texas which makes sense that they also have the 
highest jail population. Furthermore I had also calculated the total population
in Texas which came out to be 4698619 (```summary_info$total_pop_in_TX ```).
Lastly I also calculated the state and county with the highest fips  which came
out to be Weston County (```summary_info$county_highest_fips  ```) and Wyoming 
(```summary_info$state_highest_fips ```). This makes sense because Weston County
is a county in Wyoming.

```{r echo=FALSE}
summary_info
```

## Trends Chart
This chart below shows the top 10 unique counties in Texas every 5 years 
between 1990 and 2010. This shows us the black jail population over time 
throughout the years and as we can see from the chart Harris County and
Dallas County had the highest black jail population between 1990 and 2010. You
can see there is a huge difference between those two and the rest of the
counties.

```{r black_jail_pop, echo = FALSE}
ggplot(df, aes(x = Year, y = black_jail_pop)) + 
  geom_line(aes(color = County), size = 1) +
  theme_minimal()+
  labs(x = "year", y = "black Jail Population",
       title = "Black Jail Population in Texas between 1990 and 2010 ")
```


## Variable comparison Chart
The comparison chart below shows the black and white population of 10 counties
in Texas to shows the difference between the two races. As you can see the 
white population in each of these counties is significantly bigger than the
black population. I included this chart to show maybe since the black 
population is very low then the black jail population might be low as well.

```{r pop_comparison, echo = FALSE}
ggplot(dfm,aes(x = population, y = Counties)) + 
  geom_bar(aes(fill = Race),stat = "identity",position = "dodge")+
  labs(x = "Population", y = "Counties",
     title = "Black Vs white population from 10 counties in Texas")
```


## Map
The map below shows us the percentage of natives in each county of Texas. The 
places with lighter blue shows that the percentage of natives in those counties
are higher compared to the places that are more darker. The darker places show
the percentage of native population is very low. Throughout the entire map it
shows the percentage of native population spread apart in the state of Texas.

```{r perc_native_15to64, echo = FALSE}
# Percentage of native population in each county of Texas
 ggplot(map_data) +
  geom_polygon(
    mapping = aes(x = long, y = lat, group = group, fill = perc_native_15to64),
    color = "grey", size = 1.2
  ) +
  coord_map() +
  scale_fill_continuous(limits = c(0, max(map_data$perc_native_15to64)), na.value = "white") +
  blank_theme +
  ggtitle("Populations per county in TX")
```
